Wasted Money? Changes in Toronto's Annual Police Budget and Reported

### 2024-09-27 paper/paper.qmd
Styled using styler

# present the budget variables table
summary_table_bud %>%
  kable("latex", caption = "Total Police Budget Over Years {#tbl-bud}",
        col.names = c("Year", "Yearly Budget", 
                              "Change From Last", "Percent Change")) %>%
  kable_styling(latex_options = c("striped", "hold_position"))

```

@@

# present the crime variables table
summary_table_cri %>%
  kable("latex", caption = "Total Crime Reports Over Years {#tbl-cri}",
        col.names = c("Year", "Yearly Crimes", 
                              "Cleared Crimes", "Percent Cleared")) %>%
  kable_styling(latex_options = c("striped", "hold_position"))

```

@@

# take billions for budget to make a clearer graph
summary_table_bud <- summary_table_bud %>% mutate(
  bud_billion = Yearly_Budget/1000000000
)

# graphs to show change in police budget and crime rates over time
ggplot(data = summary_table_bud, aes(x = Year, y = bud_billion)) + 
  geom_bar(stat = "identity", fill = "skyblue") + 
  labs(
    x = "Year",
    y = "Yearly Budget (Billion$)"
  )

```
\vspace{1em}
\FloatBarrier

@@

# plot for crime and cleared crime count over time
ggplot(summary_table_cri, aes(x = factor(Year))) + 
  # Plot total crime count as a full bar
  geom_bar(aes(y = Yearly_Crimes, fill = "Total Crimes"), stat = "identity",
           position = "stack") +

  # Overlay cleared crimes as a portion of the total crimes
  geom_bar(aes(y = Cleared_Crimes, fill = "Cleared Crimes"), 
           stat = "identity", position = "stack") +

  # Custom colors
  scale_fill_manual(values = c("Total Crimes" = "tomato", 
                               "Cleared Crimes" = "skyblue")) +

  # Labels and theme
  labs(
    x = "Year",
    y = "Number of Crimes",
    fill = "Crime Type"
  ) +
  theme_minimal()

```

@@

# take billions for budget to make a clearer graph
combined <- combined %>% mutate(
  bud_billion = Yearly_Budget/1000000000
)

# graph
ggplot(data = combined, aes(x = bud_billion, y = Yearly_Crimes)) + 
  geom_line(stat = "identity", color = "skyblue") + 
  geom_text(aes(label = Year), 
            vjust = -0.35,
            hjust = -0.2,
            size = 2.4,
            color = "black") + 
  scale_x_continuous(limits = c(1.19, 2.5)) +
  scale_y_continuous(limits = c(118000, 173000)) +
  labs(
    x = "Yearly Police Budget (Billion$)",
    y = "Total Reported Crimes"
  ) +
  geom_point()

```

@@

# graph cleared crimes over police budget
ggplot(data = combined, aes(x = bud_billion, y = Percentage_Cleared)) + 
  geom_line(stat = "identity", color = "skyblue") + 
  geom_text(aes(label = Year), 
            vjust = -0.35,
            hjust = -0.2,
            size = 2.4,
            color = "black") + 
  scale_x_continuous(limits = c(min(combined$bud_billion), max(combined$bud_billion) + 0.1)) +
  labs(
    x = "Yearly Police Budget (Billion$)",
    y = "Percentage of Total Crimes Cleared"
  ) +
  geom_point()

```
\FloatBarrier

@@

# make a summary table
raw_bud_summary <- raw_bud %>%
  summarise(
    across(
      c(Fiscal_Year, Spending),
      list(
        Mean = ~round(mean(.), 2),
        Median = ~round(median(.), 2),
        SD = ~round(sd(.), 2),
        Min = ~round(min(.), 2),
        Max = ~round(max(.), 2)
      ),
      .names = "{col}_{fn}"
    )
  )

@@

# present the table using kable
kable(bud_summary_df_t,
      caption = "Summary Statistics for Raw Police Budget Dataset {#tbl-raw-bud}",
      col.names = c("Statistic", "Value"),
      format = "latex") %>%
  kable_styling(latex_options = c("striped", "hold_position"))

```

@@

raw_cri_summary <- raw_cri %>%
  summarise(
    across(
      c(REPORT_YEAR, CRIME_COUNT, COUNT_CLEARED),
      list(
        Mean = ~round(mean(.), 2),
        Median = ~round(median(.), 2),
        SD = ~round(sd(.), 2),
        Min = ~round(min(.), 2),
        Max = ~round(max(.), 2)
      ),
      .names = "{col}_{fn}"
    )
  )

@@

kable(cri_summary_df_t,
      caption = "Summary Statistics for Raw Crime Dataset {#tbl-raw-cri}",
      col.names = c("Statistic", "Value"),
      format = "latex") %>%
  kable_styling(latex_options = c("striped", "hold_position"))

```

>>>>

# present the budget variables table
summary_table_bud %>%
  kable("latex",
    caption = "Total Police Budget Over Years {#tbl-bud}",
    col.names = c(
      "Year", "Yearly Budget",
      "Change From Last", "Percent Change"
    )
  ) %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

@@

# present the crime variables table
summary_table_cri %>%
  kable("latex",
    caption = "Total Crime Reports Over Years {#tbl-cri}",
    col.names = c(
      "Year", "Yearly Crimes",
      "Cleared Crimes", "Percent Cleared"
    )
  ) %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

@@

# take billions for budget to make a clearer graph
summary_table_bud <- summary_table_bud %>% mutate(
  bud_billion = Yearly_Budget / 1000000000
)

# graphs to show change in police budget and crime rates over time
ggplot(data = summary_table_bud, aes(x = Year, y = bud_billion)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(
    x = "Year",
    y = "Yearly Budget (Billion$)"
  )
```
\vspace{1em}
\FloatBarrier

@@

# plot for crime and cleared crime count over time
ggplot(summary_table_cri, aes(x = factor(Year))) +
  # Plot total crime count as a full bar
  geom_bar(aes(y = Yearly_Crimes, fill = "Total Crimes"),
    stat = "identity",
    position = "stack"
  ) +

  # Overlay cleared crimes as a portion of the total crimes
  geom_bar(aes(y = Cleared_Crimes, fill = "Cleared Crimes"),
    stat = "identity", position = "stack"
  ) +

  # Custom colors
  scale_fill_manual(values = c(
    "Total Crimes" = "tomato",
    "Cleared Crimes" = "skyblue"
  )) +

  # Labels and theme
  labs(
    x = "Year",
    y = "Number of Crimes",
    fill = "Crime Type"
  ) +
  theme_minimal()
```

@@

# take billions for budget to make a clearer graph
combined <- combined %>% mutate(
  bud_billion = Yearly_Budget / 1000000000
)

# graph
ggplot(data = combined, aes(x = bud_billion, y = Yearly_Crimes)) +
  geom_line(stat = "identity", color = "skyblue") +
  geom_text(aes(label = Year),
    vjust = -0.35,
    hjust = -0.2,
    size = 2.4,
    color = "black"
  ) +
  scale_x_continuous(limits = c(1.19, 2.5)) +
  scale_y_continuous(limits = c(118000, 173000)) +
  labs(
    x = "Yearly Police Budget (Billion$)",
    y = "Total Reported Crimes"
  ) +
  geom_point()
```

@@

# graph cleared crimes over police budget
ggplot(data = combined, aes(x = bud_billion, y = Percentage_Cleared)) +
  geom_line(stat = "identity", color = "skyblue") +
  geom_text(aes(label = Year),
    vjust = -0.35,
    hjust = -0.2,
    size = 2.4,
    color = "black"
  ) +
  scale_x_continuous(limits = c(min(combined$bud_billion), max(combined$bud_billion) + 0.1)) +
  labs(
    x = "Yearly Police Budget (Billion$)",
    y = "Percentage of Total Crimes Cleared"
  ) +
  geom_point()
```
\FloatBarrier

@@

# make a summary table
raw_bud_summary <- raw_bud %>%
  summarise(
    across(
      c(Fiscal_Year, Spending),
      list(
        Mean = ~ round(mean(.), 2),
        Median = ~ round(median(.), 2),
        SD = ~ round(sd(.), 2),
        Min = ~ round(min(.), 2),
        Max = ~ round(max(.), 2)
      ),
      .names = "{col}_{fn}"
    )
  )

@@

# present the table using kable
kable(bud_summary_df_t,
  caption = "Summary Statistics for Raw Police Budget Dataset {#tbl-raw-bud}",
  col.names = c("Statistic", "Value"),
  format = "latex"
) %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

@@

raw_cri_summary <- raw_cri %>%
  summarise(
    across(
      c(REPORT_YEAR, CRIME_COUNT, COUNT_CLEARED),
      list(
        Mean = ~ round(mean(.), 2),
        Median = ~ round(median(.), 2),
        SD = ~ round(sd(.), 2),
        Min = ~ round(min(.), 2),
        Max = ~ round(max(.), 2)
      ),
      .names = "{col}_{fn}"
    )
  )

@@

kable(cri_summary_df_t,
  caption = "Summary Statistics for Raw Crime Dataset {#tbl-raw-cri}",
  col.names = c("Statistic", "Value"),
  format = "latex"
) %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```
